Single molecule orientation determination with multiview polarized illumination: modeling and microscope design

### 2017-08-02 paper/paper/paper.tex
Paper draft to Patrick.

\documentclass[10pt]{article}
\usepackage[OE]{express}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}

\begin{document}
\title{Polarized Illumination For Determining Single-Molecule Orientation:
  Modeling and Design Comparison}

@@

%%%%%%%%%%%%%%%%%%% abstract and OCIS codes %%%%%%%%%%%%%%%%
\begin{abstract*}
Abstract here. 
\end{abstract*}

\ocis{(180.0180) Microscopy; (260.5430) Polarization; (110.0110) Imaging systems; (180.2520) Fluorescence microscopy; (180.6900) Three-dimensional microscopy.}

%%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{99}
\bibitem{gallo99} K. Gallo and G. Assanto, ``All-optical diode based on second-harmonic generation in an asymmetric waveguide,'' \josab {\bfseries 16}(2), 267--269 (1999).
\end{thebibliography}

\section{Introduction}
Single molecule orientation determination is important because

A variety of techniques have been tested to determine the orientation of
single molecules:

polarized detection [Fourkas], shape analysis [Moerner], other things from
Moerner review.

We propose polarized illumination as an complementary method to polarized
detection. We get the speed (no shape analysis required). We don't waste any
dose (polarized light throws away light dose).

In this work we make three contributions:
\begin{itemize}
\item we develop the polarized illumination forward model for a broad class of
  microscopes (section 2.X-2.X)
\item we develop metrics to compare the performance of polarized illumination
  microscopes 
\item we use these metrics to compare microscopes design and make recommendations
  for polarized illumination microscopes. 
\end{itemize}

\section{Methods}
\subsection{Illumination Model}
\begin{align}
  \eta_{\text{det},\phi_{\text{det}}}  = \frac{\int_{\Omega}d\mh{r}\ \left|\mh{P}_{\text{det}}\cdot\tilde{\mb{R}}(\mh{r})(\mh{r}\times\hat{\bs{\mu}}_{\text{em}}\times\mh{r})\right|^2}{\int_{\mathbb{S}^2}d\mh{r}\left|\mh{r}\times\hat{\bs{\mu}}_{\text{em}}\times\mh{r}\right|^2}\label{eq:final}
\end{align}
where
\begin{align}
  \hat{\mb{r}} &= \sin\theta\cos\phi\hat{\mb{i}} + \sin\theta\sin\phi\hat{\mb{j}} + \cos\theta\hat{\mb{k}}\label{eq:r_coords}\ \ \ \text{is the dummy integration vector,}\\
  \hat{\bs{\mu}}_{\text{em}} &= \sin\Theta\cos\Phi\hat{\mb{i}} + \sin\Theta\sin\Phi\hat{\mb{j}} + \cos\Theta\hat{\mb{k}}\ \ \ \text{is the emission dipole moment,}\label{eq:mu_coords}\\
  \tilde{\mb{R}}(\mh{r}) &= \begin{bmatrix} \cos\theta\cos^2\phi + \sin^2\phi & (\cos\theta -1)\sin\phi\cos\phi & -\sin\theta\cos\phi\\ (\cos\theta - 1)\sin\phi\cos\phi & \cos\theta\sin^2\phi + \cos^2\phi & -\sin\theta\sin\phi \\ \sin\theta\cos\phi& \sin\theta\sin\phi & \cos\theta \end{bmatrix}\label{eq:matrix}\\&\text{is the rotation matrix that models the objective lens,}\\
  \hat{\mb{P}}_{\text{det}} &= \cos\phi_{\text{det}}\hat{\mb{i}} + \sin\phi_{\text{det}}\hat{\mb{j}}\ \ \ \text{is the pass axis of the polarizer,}\\
  \Omega &= \{(\phi, \theta)|\phi \in (0,2\pi]\ \text{and}\ \theta\in (0, \alpha]\}\ \ \ \text{is the cap of the sphere collected by the objective.}
\end{align}

\begin{align}
  \eta_{\text{exc},\phi_{\text{pol}}}(\Theta, \Phi, \alpha) &= D\{A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \cos{[2 (\Phi - \phi_{\text{pol}}})]\}.
\end{align}
\begin{subequations}
\begin{align}
  A &= \frac{1}{4} - \frac{3}{8} \cos{\alpha } + \frac{1}{8} \cos^{3}{\alpha }\\
  B &= \frac{3}{16} \cos{\alpha } - \frac{3}{16} \cos^{3}{\alpha }\\
  C &= \frac{7}{32} - \frac{3}{32} \cos{\alpha } - \frac{3}{32} \cos^{2}{\alpha } - \frac{1}{32} \cos^{3}{\alpha}\\
  D &= \frac{4}{3(1 - \cos\alpha)}.
\end{align}%
\end{subequations}

\subsection{Detection Model}
\begin{align}
  \eta_{\text{det}}(\Theta, \alpha) = 2(A + B\sin^2\Theta)
\end{align}

\subsection{Complete Forward Model}
\begin{align}
  I = I_{\text{tot}}\eta_{\text{exc}}\eta_{\text{det}}
\end{align}

\begin{figure}[htbp]
\centering\includegraphics[width=\textwidth]{single-frame}
\caption{Representative examples of the forward model for single frame
  microscopes.\newline \newline \textbf{Columns left to right:} (1) schematics of
  single frame microscopes where the solid line encloses the illumination
  solid angle, the dashed line encloses the detection solid angle, and the arrow
  indicates the transmission axis of the illumination polarizer; (2) the excitation
  efficiency see equation XX; (3) the detection efficiency see equation X; (4)
  the total efficiency see equation X.\newline \newline \textbf{Rows top to
    bottom:} (1) epi-illumination (NA = 1.1) with $x$-polarized light and
  epi-detection (NA = 1.1); (2) orthogonal illumination (NA = 0.8) and detection
  (NA = 0.8); (3) oblique illumination (NA = 0.8) and detection (NA = 0.8).}
\end{figure}

\subsection{Microscope Designs}
Widefield illumination vs light sheet illumination.

- Light sheet has low illumination. 
- Restricted to orthogonal detection.

\subsection{Evaluation Metrics}
\begin{align}
  F &= \sum_{i=0}^N \frac{1}{I}
  \begin{bmatrix}
    \frac{\partial I}{\partial \Theta}\frac{\partial I}{\partial \Theta}&\frac{\partial I}{\partial \Theta}\frac{\partial I}{\partial \Phi}\\\\
    \frac{\partial I}{\partial \Theta}\frac{\partial I}{\partial \Phi}&\frac{\partial I}{\partial \Phi}\frac{\partial I}{\partial \Phi}\\    
  \end{bmatrix}\\
  \sigma_{\Omega} &= \frac{\sin\Theta}{\sqrt{\text{det}\{F\}}}
\end{align}
We call $\sigma_{\Omega}$ the \emph{solid-angle uncertainty}. 
We sampled $\sigma_{\Omega}$ on 100,000 approximately equally
spaced points on the unit sphere and calculated the \emph{coefficient of variation}
$c_v$ of these points where
\begin{align}
  c_v = \frac{\sigma_{\sigma_{\Omega}}}{\mu_{\sigma_{\Omega}}}
\end{align}

\section{Results}
\subsection{One-Arm Designs}
\begin{figure}[htbp]
\centering\includegraphics[width=\textwidth]{single-arm}
\caption{\textbf{Left:} Schematic of a single-arm four-frame epi-illumination
  microscope with NA = 0.8 and $n = 1.33$.  \textbf{Right:} Solid
  angle uncertainty for the same microscope when $I_{\text{tot}} = 1000$
  photons. Note the regions of high uncertainty near the equator and poles
  as noted in XXX.}
\end{figure}

Can't use $c_v$ because it doesn't converge.

High variance around equator and poles.

\subsection{Two-Arm Designs}
Widefield vs light sheet.

Widefield NA vs angle/.

Sweep asymmetry for full NA light sheet.

\subsection{Three-Arm Designs}

\section{Discussion}

\section{Conclusion}

>>>>

\documentclass[10pt]{article}
\usepackage[OE]{express}
\usepackage{float}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}

\begin{document}
\title{Single molecule orientation determination with multiview polarized
  illumination: modeling and microscope design}

@@

%%%%%%%%%%%%%%%%%%% abstract and OCIS codes %%%%%%%%%%%%%%%%
\begin{abstract*}
  We investigate the use of polarized illumination in multiview microscopes for
  determining the orientation of single molecules. First, we relate the
  orientation of single molecules to measurable intensities in multiview
  microscopes and develop an information theoretic metric---the solid angle
  uncertainty---to compare the ability of multiview microscopes to estimate the
  orientation of single molecules. Next, we compare a broad class of microscopes
  using this metric---single- and dual-view microscopes with varying
  illumination polarization, illumination numerical aperture (NA), detection NA,
  obliquity, asymmetry, and exposure. We find that multiview microscopes are
  better at determining the orientation of single molecules than single-view
  microscopes. We also find that choosing a small illumination NA and a large
  detection NA are good design choices, multiview widefield microscopes can
  benefit from oblique illumination and detection, and asymmetric NA microscopes
  can benefit from exposure asymmetry.
\end{abstract*}

\ocis{(110.0110) Imaging systems; (180.2520) Fluorescence microscopy; (180.0180)
  Microscopy; (180.6900) Three-dimensional microscopy; (130.5440)
  Polarization-selective devices; (260.5430) Polarization.}

%%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%%
%% Comment before submission
\bibliography{paper}
\bibliographystyle{osajnl}

%% Copy .bbl and uncomment these before submission
% \begin{thebibliography}{99}
% \bibitem{gallo99} K. Gallo and G. Assanto, ``All-optical diode based on second-harmonic generation in an asymmetric waveguide,'' \josab {\bfseries 16}(2), 267--269 (1999).
% \end{thebibliography}

\section{Introduction}
The orientation of single molecules is a valuable reporter for biological
processes. By imaging biological samples with fluorescent probes that are fixed
to known structures, researchers have studied the dynamics of motor proteins
\cite{peterman2001, forkey2003}, DNA \cite{backer2016}, actin \cite{mehta2016},
septin \cite{demay2011, mcquilken2017}, and membranes \cite{anantharam2010}. The
number of techniques available to biologists for measuring the orientation of
single molecules in live cells is continually expanding. We follow a recent
review \cite{backlund2014} and compare three major categories of single molecule
orientation determination techniques:

\textbf{Spatial variation of emission} techniques image single molecules and use
the spatial distribution of intensity in the back focal plane \cite{lieb2004} or
in the image plane \cite{backer2014} to estimate their orientation. The best
techniques in this category use defocusing or phase masks in the back focal
plane to encode orientation and position information in the image
\cite{agrawal2012}. These techniques can estimate the orientation of
molecules in all orientations and only require a single frame, but they
require complex reconstruction algorithms, expensive high numerical aperture
(NA) optics, and sensitive detectors.

\textbf{Spatial variation of illumination polarization} techniques vary the
polarization and intensity in the illumination path to excite molecules in
specific orientations in the focal volume \cite{debarre2004}. Although these
techniques can accurately estimate the orientation of molecules in all
orientations, they require scanning which makes these techniques too slow for
live cell imaging.

\textbf{Polarized illumination or detection} techniques exploit the anisotropic
absorption and emission pattern of single molecules to estimate their
orientation. By measuring the total intensity in the detector plane during
several polarization orientations on the illumination or detection path, the
orientation of the molecule can be estimated. These techniques are easy to
implement---changing the illumination or detection polarization is simple with a
polarization splitter \cite{mehta2016} or universal compensator
\cite{oldenbourg1995}---and the reconstruction methods are straightforward
\cite{fourkas2001, mehta2016, backer2016}. The main drawback is speed---these
techniques require the collection of several polarization frames to find the
orientation of a single molecule. Despite the requirement for multiple frames,
polarized illumination is the method of choice for live cell imaging because of
its good combination of speed and ease of implementation. For these reasons,
polarized illumination is the method we consider and extend in this paper.

All single molecule orientation determination methods suffer from some degree of
anisotropic orientation uncertainty, i.e., molecules in some orientations
cannot be oriented as well as molecules in others. In some cases the
orientation cannot be determined at all---the forward model can be degenerate
for specific dipole orientations \cite{fourkas2001, lu2008}. We feel that the
importance of isotropic orientation uncertainty has been underappreciated. To
our knowledge, the only authors who have considered anisotropic orientation
uncertainty have only analyzed a small set of molecule orientations intead of
all orientations \cite{agrawal2012}. Our view is that an ideal technique for
determining the orientation of single molecules can reconstruct the
orientation with a small and nearly uniform uncertainty for all molecule
orientations.

Recently, there has been increasing interest in multiview microscopy techniques
for biological imaging \cite{wu2013, keller2015, wu2016, wu2017}. Multiview
microscopes address two key limitations of single view microscopes. First,
multiview microscopes can achieve nearly isotropic resolution compared to the
poor axial resolution of single view microscopes. Second, multiview microscopes
can easily incorporate selective illumination to reduce phototoxicity. For
example, compare light-sheet illumination to wide-field illumination where
most of the illumination light excites out of focus molecules. These
advantages make multiview microscopes good candidates for imaging live
biological specimens for long periods.

In this work we explore the use of polarized illumination in multiview
microscopes for determining the orientation of single molecules. Existing
multiview microscopes can easily be outfitted with fast-switching polarizers
that do not degrade image quality, so polarized illumination is a natural way to
augment multiview microscopes for measuring the orientation of single
molecules. Furthermore, multiview microscopes can achieve nearly isotropic
resolution while delivering selective illumination. We hypothesize that
multiview microscopes will also provide a small and uniform orientation
uncertainty for single molecules in all orientations. In section \ref{methods}
we develop the required theory for polarized illumination microscopy and develop
information theoretic metrics to compare multiview microscopes. In section
\ref{results} we compare the results for a wide range of multiview microscope
designs. Finally, in section \ref{discussion} we discuss the results and their
impact on polarized multiview microscope design.

\section{Methods}\label{methods}
In this section we develop a method to compare multiview polarized illumination
microscopes for the task of reconstructing molecular orientations. In sections
\ref{excitation}--\ref{forward} we develop the forward model for single-frame
microscopes with a fixed illumination and detection configuration. In section
\ref{designs} we list a variety of ways that single-frame microscopes can be
combined to create experimentally realizable multi-frame microscopes. Finally,
in section \ref{metrics} we develop metrics that we will use to compare
multi-frame microscopes.

\subsection{Absorption Efficiency}\label{excitation}
Consider polarized widefield K\"ohler illumination incident on a single
molecule at the focal point with the optical axis along the $\mh{z}$ axis. In
this section we will calculate the \emph{absorption efficiency} in this
geometry---the fraction of the incident power that is absorbed by the molecule.

We start with the unit electric field transmitted by the polarizer,
\begin{align}
 \hat{\mb{E}}_{\text{exc}} = \cos\phi_{\text{exc}}\hat{\mb{x}} +
\sin\phi_{\text{exc}}\hat{\mb{y}},\label{eq:excfield}
\end{align}
where $\phi_{\text{exc}}$ is the angle between the transmission axis of the
excitation polarizer and the $\mh{x}$ axis. Next, we model the action of the an
ideal, aplanatic, polarization preserving illumination objective by multiplying
the electric field vector by a direction dependent rotation matrix,
  \begin{align}
  \mb{R}(\mh{r}) &= \begin{bmatrix} \cos\theta\cos^2\phi + \sin^2\phi & (\cos\theta -1)\sin\phi\cos\phi & -\sin\theta\cos\phi\\ (\cos\theta - 1)\sin\phi\cos\phi & \cos\theta\sin^2\phi + \cos^2\phi & -\sin\theta\sin\phi \\ \sin\theta\cos\phi& \sin\theta\sin\phi & \cos\theta \end{bmatrix}
  \end{align}
  where
  \begin{align}
   \hat{\mb{r}} = \sin\theta\cos\phi\hat{\mb{x}} + \sin\theta\sin\phi\hat{\mb{y}} + \cos\theta\hat{\mb{z}} 
  \end{align}
  is a unit vector that points from the origin to the point where the incident
  electric field intersects the Gaussian sphere of the illumination
  objective. Next, we find the power that excites the molecule by taking the
  dot product of the rotated incident field $\mb{R}(\mh{r})\mh{E}_{\text{exc}}$
  and the absorption dipole moment
\begin{align}
  \hat{\bs{\mu}}_{\text{abs}} = \sin\Theta\cos\Phi\hat{\mb{x}} + \sin\Theta\sin\Phi\hat{\mb{y}} + \cos\Theta\hat{\mb{z}}
\end{align}
then taking its modulus squared. If the incident fields are spatially incoherent
then we can find the total power that excites the molecule by integrating over the
illuminated region of the Gaussian sphere $\Omega$ where
\begin{align}
  \Omega &= \{\phi, \theta\ |\ 0 < \phi \leq 2\pi, 0 < \theta \leq \alpha\}, \\
  \alpha &= \arcsin\left(\frac{\text{NA}}{n}\right), \label{eq:alpha}
\end{align}
NA is the illumination numerical aperture, and $n$ is the index of refraction of
the medium.

Finally, we find the fraction of the incident power absorbed by the molecule
by dividing the total absorbed power by the total incident power. The complete
expression for the absorption efficiency in vector notation is
\begin{align}
  \eta_{\text{abs}} = \frac{\int_{\Omega}d\mh{r}|\hat{\bs{\mu}}_{\text{abs}}\cdot\mb{R}(\mh{r})\mh{E}_{\text{exc}}|^2}{\int_{\Omega}d\mh{r}}\label{eq:abs}. 
\end{align}

We substitute equations \ref{eq:excfield}--\ref{eq:alpha} into equation
\ref{eq:abs}, evaluate the integrals, and simplify to express the absorption
efficiency in scalar notation as
\begin{align}
  \eta_{\text{abs}} &= D\{A + B\sin^{2}{\Theta} + C\sin^{2}{\Theta} \cos{[2 (\Phi - \phi_{\text{exc}}})]\}\label{eq:scalarabs}
\end{align}
where
\begin{subequations}
\begin{align}
  A &= \frac{1}{4} - \frac{3}{8} \cos{\alpha } + \frac{1}{8} \cos^{3}{\alpha }\\
  B &= \frac{3}{16} \cos{\alpha } - \frac{3}{16} \cos^{3}{\alpha }\\
  C &= \frac{7}{32} - \frac{3}{32} \cos{\alpha } - \frac{3}{32} \cos^{2}{\alpha } - \frac{1}{32} \cos^{3}{\alpha}\\
  D &= \frac{4}{3(1 - \cos\alpha)}.
\end{align}\label{eq:coefficients}%
\end{subequations}

At the beginning of this section we assumed that the molecule was located at
the focal point of the condenser. If the illumination objective is aplanatic
then we can relax this assumption and the above expressions are valid for
molecules anywhere in the focal plane.

We also assumed that we used incoherent K\"ohler illumination. If we illuminate
the back aperture with a weakly focused laser beam and scan the laser beam
slowly---if the scan velocity is much less than the beam waist divided by the
coherence time---we can ignore longitudinal excitation and the molecule will
be excited as if a single plane was incident. Therefore, we can find the
absorption efficiency of a molecule under weakly focused scanned laser
illumination by taking the limit of equation \ref{eq:scalarabs} as
$\alpha \rightarrow 0$ giving
\begin{align}
  \eta_{\text{abs}} &= \sin^2\Theta\cos^2(\Phi - \phi_{\text{exc}})
\end{align}
which is recognizable as Malus' law generalized to three dimensions. Notice that
the absorption efficiency can take its a maximum range of values
$\eta_{\text{abs}} \in [0, 1]$ when $\alpha=0$.

\subsection{Detection Efficiency}\label{detection}
Fourkas calculated the detection efficiency of a single molecule when an
objective with a polarizer is aligned along the $\mh{z}$ axis
\cite{fourkas2001}. We use his expressions to calculate the detection efficiency
without a polarizer as
\begin{align}
  \eta_{\text{det}} = 2(A + B\sin^2\Theta). \label{eq:scalardet}
\end{align}
The detection efficiency only depends on $\Theta$, not $\Phi$, because there is
no detection polarizer. We have assumed that the absorption dipole moment is
equal to the emission dipole moment.

Note that $A$ and $B$ in equation \ref{eq:coefficients} are a factor of
$\frac{3}{2}$ larger than the expressions given by Fourkas. We found that
Fourkas' expressions were incorrectly normalized (the limit of
$\eta_{\text{det}}$ as $\alpha\rightarrow \pi/2$ should be 1), and the extra
factor of $\frac{3}{2}$ corrects the error. This correction does not affect
Fourkas' orientation reconstruction but it does mean that Fourkas' algorithm
under predicts the total emitted intensity by a factor of $\frac{3}{2}$.

Also note that the detection efficiency can take its a maximum range of values
when $B$ takes its maximum value, i.e. when
$\alpha=\arccos\left(\frac{1}{\sqrt{3}}\right) \approx 54.7^{\circ}$. When $n = 1.33$,
this condition corresponds to $\text{NA} \approx 1.1$.

\subsection{Oblique Illumination and Detection}\label{oblique}
In sections \ref{excitation} and \ref{detection} we assumed that both the
illumination and detection objectives had $\mh{z}$ aligned optical axes. To extend
the forward model to oblique optical axes, we express the dipole orientation in
rotated coordinates using the following expressions
\begin{align}
    \Theta' &= \arccos\left(\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta\right)\label{eq:thetap}\\
  \Phi' &=
          \begin{cases}
            \arccos\left(\dfrac{\cos\psi\cos\Phi\sin\Theta - \sin\psi\cos\Theta}{\sqrt{1 - (\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta)^2}}\right) \ \ \ &0 \leq \Phi < \pi  \\
            -\arccos\left(\dfrac{\cos\psi\cos\Phi\sin\Theta - \sin\psi\cos\Theta}{\sqrt{1 - (\sin\psi\cos\Phi\sin\Theta + \cos\psi\cos\Theta)^2}}\right) \ \ \ &-\pi \leq \Phi < 0\\
          \end{cases}
\end{align}
where $\psi$ is the angle of a right handed rotation about the $\mh{y}$ that maps the $\mh{z}$ axis onto the new optical axis.

\subsection{Single-Frame Microscopes}\label{forward}
The detected intensity is the product of the the sample exposure
$I_{\text{tot}}$, the absorption efficiency, and the detection efficiency
\begin{align}
  I = I_{\text{tot}}\eta_{\text{abs}}\eta_{\text{det}}\label{eq:forward_frame}.
\end{align}
Equation \ref{eq:forward_frame} is the forward model for a \emph{single-frame
  microscope}. Figure \ref{fig:single-frame} shows the efficiencies from three
representative examples of single-frame microscopes.

The intensity measured by a single-frame microscope does not give us enough
information to reconstruct the orientation of a single molecule. Next, we
will consider combining several single-frame microscopes to create
\emph{multi-frame microscopes.}

Note our distinction between \emph{frames} and \emph{views}. Frames refer to
single intensity measurements with fixed illumination, detection, and
polarization geometries. Views refers to the detection objectives, and multiview
microscopes use multiple detection objectives.

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{single-frame}
\caption{Representative examples of single-frame microscopes.\newline \newline
  \textbf{Columns left to right:} 1) schematics of single-frame microscopes
  where the solid line encloses the illumination solid angle, the dashed line
  encloses the detection solid angle, and the arrow indicates the transmission
  axis of the illumination polarizer; 2) the absorption efficiency, see equation
  \ref{eq:scalarabs}; 3) the detection efficiency, see equation
  \ref{eq:scalardet}; (4) the total efficiency, the product of the absorption and
  detection efficiencies.\newline \newline \textbf{Rows top to bottom:} 1)
  epi-illumination ($\text{NA} = 1.1$) with $x$-polarized light and
  epi-detection (NA = 1.1); 2) orthogonal illumination ($\text{NA} = 0.8$) and
  detection ($\text{NA} = 0.8$); 3) oblique illumination ($\text{NA} = 0.8$) and
  detection ($\text{NA} = 0.8$). All simulations use $n=1.33$.}
  \label{fig:single-frame}
\end{figure}

\subsection{Multi-Frame Microscopes}\label{designs}
One way to collect multiple frames is to add a universal compensator to the
illumination arm and rapidly select the incident polarization by changing
$\phi_{\text{exc}}$ \cite{oldenbourg1995}. All of the multi-frame microscopes we
will consider in this paper use four frames per view with the illumination
polarization separated by 45${}^{\circ}$.

We also consider several dual-view designs. We evaluate two-view designs that
allow for illumination and detection from both objectives (see \cite{wu2013} for
a dual-view design). In all cases we consider the effect of varying the
illumination and detection numerical aperture. We also consider asymmetric NA
designs (see \cite{wu2017} for an asymmetric NA design) and the effect of
sample exposure.

All multiview microscopes are subject to steric constraints---the objectives
must not collide with the cover slip or each other. We only consider microscope
designs that meet these criteria.

\subsection{Evaluation Metrics}\label{metrics}
Our goal is to evaluate the ability of a microscope design to estimate the
orientation of a single molecule (via parameters $\Theta$ and $\Phi$) from
intensity data.

For each microscope design and molecule orientation we calculate the Fisher
information matrix. If the intensity is Poisson distributed the
Fisher information matrix is given by
\begin{align}
  \mb{F} &= \sum_{k=1}^N \frac{1}{I_k}
  \begin{bmatrix}
    \dfrac{\partial I_k}{\partial \Theta}\dfrac{\partial I_k}{\partial \Theta}&\dfrac{\partial I_k}{\partial \Theta}\dfrac{\partial I_k}{\partial \Phi}\\\\
    \dfrac{\partial I_k}{\partial \Theta}\dfrac{\partial I_k}{\partial \Phi}&\dfrac{\partial I_k}{\partial \Phi}\dfrac{\partial I_k}{\partial \Phi}\\    
  \end{bmatrix}
\end{align}
where $I_k$ is the intensity measured in the $k$th frame of an $N$ frame
microscope.

A common way to evaluate the ability to estimate the parameters $\Theta$ and
$\Phi$ from the data is to calculate the Cramer-Rao lower bound (CRLB) for each
parameter \cite{kay1993}. The CRLBs are given by the diagonal elements of the
inverse of the Fisher information matrix, and they give the minimum variance of
an unbiased estimator for each parameter. Agrawal et. al. used the square root
of the product of the CRLBs multiplied by the Jacobian determinant,
$\sin\Theta\sqrt{\mb{F}^{-1}_{0,0}\mb{F}^{-1}_{1,1}}$, to find the area of
uncertainty in parameter space \cite{agrawal2012}.

CRLBs and the associated area of uncertainty are parametrization dependent,
i.e., if we choose a different coordinate system these metrics will change. We
would like to compare microscope designs without choosing a parametrization, so
instead we use
\begin{align}
  \sigma_{\Omega} &\equiv \sin\Theta\sqrt{\text{det}\{\mb{F}^{-1}\}}
\end{align}
as our evaluation metric. We take the determinant of the inverse Fisher
information matrix---a parametrization independent value called the
\emph{generalized variance} \cite{anderson1958}---take the square root, then
multiply by the Jacobian determinant, $\sin\Theta$. We call $\sigma_{\Omega}$
the \emph{solid-angle uncertainty} because it has units of steradians and is a
parametrization independent measure of the orientation uncertainty.

For each multi-frame microscope we calculate $\sigma_{\Omega}$ at 10,000
approximately equally spaced points on the unit sphere. A desirable microscope
design will have a small solid-angle uncertainty that is uniform for all
molecule orientations. The most straightforward way to find the location and
scale of the solid-angle uncertainty is to use the mean and variance,
respectively. Because the solid-angle uncertainty admits extremely large values
that can change the mean and variance dramatically depending on the sample
points, we use the median and median absolute deviation (MAD)---the median of
the data's absolute difference from the median---as robust replacements for the
mean and variance.

To compare multi-frame microscopes with different numbers of frames fairly, we
kept the total exposure to the sample constant by choosing
$I_{\text{tot}} = \frac{4000}{N}$. Note that $I_{\text{tot}}$ is a measure of
exposure to the sample, not the detected intensity or the total intensity
emitted by the molecule.

\section{Results}\label{results}
Figure \ref{fig:single-arm} shows our results for single-view designs where a
single objective is used for illumination and detection. We swept through the
illumination and detection NA while keeping the sample exposure constant, and we
found that the lowest median and MAD of the solid-angle uncertainty occurs with
a small illumination NA and a large detection NA where
$\alpha \approx 57.4^{\circ}$. A small illumination NA maximizes the range of
absorption efficiencies, while a large detection NA maximizes the number of
detected photons and the range of the detection efficiencies. Note the relative
importance of illumination and detection NAs---reducing the illumination NA (an
inexpensive modification) improves the solid angle uncertainty much less than
increasing the detection NA (an expensive modification). Also note from Figure
\ref{fig:single-arm}b) that single-view designs suffer from high orientation
uncertainty when the molecules are oriented along the optical axis (because the
molecules are not efficiently excited) and in the transverse plane (because it
is ambiguous whether the molecule is oriented above or below the transverse
plane).

\begin{figure}[htbp]
\centering\includegraphics[width=\textwidth]{single-arm}
\caption{Single-view microscope with varying illumination and detection
  NA. a) Schematic of a single-arm four-frame epi-illumination microscope. The
  red arrows indicate the four illumination polarization orientations---one for
  each frame. b) Solid angle uncertainty for the microscope in a). c) Median of
  the solid-angle uncertainty as a function of illumination and detection NA. d)
  MAD of solid-angle uncertainty as a function of illumination and detection
  NA. The microscope in a) and b) is indicated by a cross in c) and d).}
\label{fig:single-arm}
\end{figure}

Figures \ref{fig:symmetric-widefield} and \ref{fig:double-arm} show our results
for dual-view designs. We illuminate from one objective and detect from the
other for four polarization frames, then we repeat the polarization frames with
the same objectives in reversed roles. In Figure \ref{fig:symmetric-widefield}
we sweep through the NA of both objectives and the angle between the objectives
while considering steric constraints. Note how adding a second view removes the
high uncertainty region in the transverse plane. The dual-view microscope still
has high uncertainty regions near the optical axes, but the uncertainty is
reduced compared to the single-view microscope. We find that the lowest median
of the solid-angle uncertainty occurs with the largest possible NA
objectives. We also find that increasing NA always lowers the MAD, but
orthogonal arms are not always best. At large NA, it is advantageous to move the
objectives together or against the cover slip. Given the high uncertainty in the
transverse plane for a single view shown in Figure \ref{fig:single-arm}b), we
can see why oblique views perform better than orthogonal views---oblique views
are complementary because the regions of high orientation of uncertainty from
each view do not overlap.

\begin{figure}[htbp]
\centering\includegraphics[width=\textwidth]{symmetric-widefield}
\caption{Dual-view symmetric widefield designs with varying NA and angle between
  objectives. First we illuminate with the first objective (red solid) and
  detect from the second objective (red dashed). Then we illuminate from the
  second objective (blue solid) and detect from the first objective (blue
  dashed). See the Fig. \ref{fig:single-arm} caption for an explanation of
  b)-d).}
\label{fig:symmetric-widefield}
\end{figure}

Figure \ref{fig:double-arm} shows our results when we used a symmetric widefield
design and varied the illumination and detection NA. Our results are similar to
the single view case in Figure \ref{fig:single-arm}. We find that the best
designs use a small illumination NA and a large detection NA where
$\alpha \approx 57.4^{\circ}$. This means that dual view light sheet illumination
geometries are an excellent choice for uniformly reconstructing the orientation
of single molecules.

\begin{figure}[htbp]
\centering\includegraphics[width=\textwidth]{double-arm}
\caption{Symmetric widefield designs with varying illumination and detection NA. See the
  caption for Fig. \ref{fig:symmetric-widefield} for an explanation of a) and
  the caption for Fig. \ref{fig:single-arm} for an explanation of b)-d).}
\label{fig:double-arm}
\end{figure}

Figure \ref{fig:asymmetric-double} shows our results for asymmetric light sheet
illumination designs. We kept the two views orthogonal and used light sheet
illumination on both sides while sweeping through the NA asymmetry and the
sample exposure asymmetry while keeping the total sample exposure constant. We
found that symmetric designs are at a local minimum of the median and MAD of the
solid-angle uncertainty. We found that at extreme asymmetries where the exposure
from the low NA arm is much larger than the exposure from the high NA arm
(top-left and bottom-right corners of Figures \ref{fig:asymmetric-double}c) and
d)) the median and MAD are comparable to the symmetric light-sheet
microscope. This means that if a very high NA objective is available and a low
NA objective can provide light sheet illumination, this microscope will perform
comparably with a dual-view microscope with two objectives of medium NA. Figure
\ref{fig:asymmetric-double} also shows that designs with slightly asymmetric NA
can trade off a low solid-angle uncertainty median for a low solid-angle
uncertainty MAD by changing the sample exposure ratio.

\begin{figure}[htbp]
  \centering\includegraphics[width=\textwidth]{asymmetric-double}
  \caption{Asymmetric light-sheet illumination designs with varying NA and
    sample exposure asymmetry. See the Fig. \ref{fig:symmetric-widefield}
    caption for an explanation of a) (note that the illumination NA is small so
    the illumination is represented by a dot instead of a solid line) and the
    Fig. \ref{fig:single-arm} caption for an explanation of b)-d).}
\label{fig:asymmetric-double}
\end{figure}

% Figure \ref{fig:triple-arm} shows the results when we considered symmetric
% light-sheet designs with a lower detection arm. We swept through the NA of the
% upper arms and NA of the lower arm. The results show that the best designs have
% the largest possible detection upper and lower NA. Note to Patrick: I used the
% same efficiencies for the upper and lower arms. I'm still thinking/unsure about
% the factor of $\sim$1/9 due to rolling shutter on the lower objective. If we do
% end up using this factor it will weight the importance of a large NA on the
% upper objectives. There is a small dip in the MAD when the upper and lower
% objectives have the same NA, but it is still better to use the largest possible
% lower NA.

% \begin{figure}[htbp]
% \centering\includegraphics[width=\textwidth]{triple-arm}
% \caption{Symmetric light-sheet illumination with lower detection arm and
%     varying upper and lower NA. See the Fig. \ref{fig:single-arm} caption for
%     an explanation of a)-d).}
% \label{fig:triple-arm}
% \end{figure}

\section{Discussion and Conclusions}\label{discussion}
% Fourkas' expressions underpredict the fraction of the total power collected by
% an objective with a polarizer by a constant fraction of
% $\frac{3}{2}$. Therefore, using Fourkas' expressions to find $I_{\text{tot}}$
% will result in overpredictions of $I_{\text{tot}}$, but the incorrect
% expressions will not cause any error in the predictions of molecular
% orientation.

% Lu and Bout \cite{lu2008} study the effect of noise on the reconstruction of single fluorophores using Fourkas' expressions. Their main results are unaffected by the correction introduced in these notes, but because these notes predict a higher intensity than Fourkas, Lu and Bout overpredict the effect of noise in their model.
